Seagrass mapping and monitoring along the coasts of Crete, Greece

### 2011-01-31 MScDiplom_Lemenkova.tex
written Abstract of MSc Thesis

\section*{Abstract}
	\markright{Summary Abstract of the MSc Thesis}
	\addcontentsline{toc}{section}{Abstract}
\pagebreak

@@

in nutrient limitation of seagrass Posidonia oceanic growth in the NW Mediterranean}.

@@

marine phanerogam \textit{	extit{Posidonia oceanica}} in the Pagassitikos Gulf, Greece}. Journal of

@@

a \textit{	extit{Posidonia oceanica}} landscape and its reference for a monitoring plan}. Marine Ecology.

@@

Internet version: Available at http://www.seagrasses.org \pageref{Borum04}

@@

www.gisdevelopment.net \pageref{Bharathi03}

@@

http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf \pageref{Carter08}

@@

Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. www.sosmalaysia.org \pageref{Choo06}

@@

Dehouck A., Martiny N., Froidefond J.-M., Sénéchal N. 2008. \emph{In-Water Reflectance

@@

(\textit{Posidonia oceanica}) fish farm-induced decline}. http://arxiv.org/ftp/qbio/
papers/0611/0611006.pdf \pageref{Diaz-Almela06}

@@

http://www.unesco.org/csi/pub/source/rs.htm \pageref{Edwards00}

@@

Education and Research Foundation doi: 10.2112/07-0882.1 \pageref{Fletcher09}

@@

\textit{	extit{Posidonia oceanica}} seagrass beds in the Port-Cross National Park (north-western

@@

Wollongong. Research Online, http://ro.uow.edu.au/theses/472/ \pageref{Fyfe03}

@@

Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008 \pageref{}

@@

Possible?} http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199 Proceedings,

@@

http://www.igarss08.org/Abstracts/pdfs/3888.pdf \pageref{Matarrese06}

@@

http://www.seagrasswatch.org \pageref{McKenzieetal03}

@@

Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617, online:
http://www.idealibrary.com \pageref{Nagelkerken00}

@@

\emph{SPOT, Official Website} http://www.spotimage.com/web/en/172-spot-images.php \pageref{SPOT}

@@

ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf \pageref{Peneva07}

@@

Data. Proceedings of Asian Conference on Remote Sensing}. http://eprints.utm.my/2161/ \pageref{Rasib97}

@@

factors responsible for spatial variation}. Available on web: http://www.tesisenxarxa.net/ \pageref{Villegas06}

@@

Wabnitz C.C., Andréfouët S., Torres-Pulliza D., Müller-Karger F. E. and Kramer P.A.

>>>>

\section*{Abstract}
	\markright{Summary Abstract of the MSc Thesis}
	\addcontentsline{toc}{section}{Abstract}
The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones. The only flowering plant in the world that is able to live completely submerged, seagrasses play vital role in the marine ecosystems of the World Ocean. Seagrasses are the most important component in the environmental food chain of the coastal ecosystems, being a vital food source for various marine species (e.g. fish, dugongs, turtles, swans), and a producer of organic matter, which is the very basis of the food web.\\
The \textit{P.oceanica} seagrass is an endemic for the Mediterranean region, and a main species in the marine coastal environment of Greece. Meadows of \textit{P.oceanica} are subjected to the human activities, because they occur in coastal areas, where they are affected both by anthropogenic and by climatic and environmental factors.  Nowadays \textit{P.oceanica} is in the alarming state of regression, because of the deterioration of the environment in the Mediterranean Sea. Due to these reasons, \textit{P.oceanica} is a protected species since 1988 in some European countries (France). Monitoring \textit{P.oceanica} is therefore an important contribution to the saving and protecting the environment of Mediterranean region.\\
The current MSc thesis focuses on the monitoring of seagrass \textit{P.oceanica} along the northern coasts of Crete Island, Greece, and investigates the application of the remote sensing techniques for the seagrass mapping. 
This research was articulated in two parts, where the first one involves an ecological approach to the seagrass distribution in various regions around the globe and the experience of seagrass monitoring nowadays. The second part of this work has technical character and investigates the application of the remote sensing techniques towards seagrass mapping. It, furthermore, focuses on the optical properties of the \textit{P.oceanica} and other seafloor cover types, and studies distinguishability of various seafloor cover types. Studies of the optical characteristics of separate seafloor cover types were made with purpose to clarify, whether their spectral properties change with varying environmental conditions. \\
Special attention has been drawn on the role of environmental factors on the distribution of \textit{P.oceanica} along the coasts of Crete, and in particular, how the optical properties of the seafloor cover types, i.e. spectral reflectance, are being changed under varying external conditions, e.g. water column, amount of suspended particles and sediments in the seawater, and water temperature. For this purpose we studied differences in the spectral reflectance of \textit{P.oceanica} and other bottom cover types at three distinct depths. The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions. We applied WASI simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (MERIS, SeaWiFS, CZCS and MODIS), in order to understand their suitability and possible limitations for the seagrass mapping.\\
Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos). The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer. Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and Landsat ETM+ scenes.

@@

in nutrient limitation of seagrass \textit{Posidonia oceanica} growth in the NW Mediterranean}.

@@

marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of

@@

a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.

@@

Internet version: Available at \href{http://www.seagrasses.org}{Seagrasses.org} \pageref{Borum04}

@@

\href{www.gisdevelopment.net}{gisdevelopment} \pageref{Bharathi03}

@@

\href{http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf}{precedings.nature.com} \pageref{Carter08}

@@

Choo, C.K. 2006. \emph{Save our Seahorses}. First Edition. \href{www.sosmalaysia.org}{sosmalaysia} \pageref{Choo06}

@@

Dehouck A., Martiny N., Froidefond J.-M., S\'en\'echal N. 2008. \emph{In-Water Reflectance

@@

(\textit{Posidonia oceanica}) fish farm-induced decline}. \href{http://arxiv.org/ftp/qbio/
papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}

@@

\href{http://www.unesco.org/csi/pub/source/rs.htm}{UNESCO} \pageref{Edwards00}

@@

Education and Research Foundation, doi: 10.2112/07-0882.1 \pageref{Fletcher09}

@@

\textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western

@@

Wollongong. Research Online, \href{http://ro.uow.edu.au/theses/472/}{University of Wollongong Thesis Collection} \pageref{Fyfe03}

@@

Frederiksen M, Grau A, Karakassis I, Marbà N, Mirto S, Pérez M, Pusceddu A, Tsapakis M.2008

@@

Possible?} \href{http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199}{ieeexplore.ieee.org} Proceedings,

@@

\href{http://www.igarss08.org/Abstracts/pdfs/3888.pdf}{International Geoscience and Remote Sensing Symposium} \pageref{Matarrese06}

@@

\href{http://www.seagrasswatch.org}{seagrasswatch} \pageref{McKenzieetal03}

@@

Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617 \pageref{Nagelkerken00}

@@

\emph{SPOT, Official Website} \href{http://www.spotimage.com/web/en/172-spot-images.php}{SPOT} \pageref{SPOT}

@@

\href{ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf}{ecn.purdue.edu} \pageref{Peneva07}

@@

Data. Proceedings of Asian Conference on Remote Sensing}. \href{http://eprints.utm.my/2161/}{eprints.utm.my} \pageref{Rasib97}

@@

factors responsible for spatial variation}. Available on web: \href{http://www.tesisenxarxa.net/}{tesisenxarxa} \pageref{Villegas06}

@@

Wabnitz C.C., Andréfouët S., Torres-Pulliza D., M\"uller-Karger F. E. and Kramer P.A.
